---
title: "Module-2"
author: "Sankalp Biswal"
date: "2023-11-14"
output: html_document
---
## 1.Importing Libraries
```{r}
library(pacman)
library(ggplot2)
library(tidyverse)
library(janitor)
library(lubridate)
library(gmodels)
library(plotly)
library(reshape2)
library(dplyr)
library(purrr)
library(knitr)
library(kableExtra)
```

## 2.Importing dataset

* Selecting variables of interest
```{r}
property_dataset <- read.csv("fy2023-property-assessment-data.csv")

property_subset <- subset(property_dataset,select = c(CITY,LU,LU_DESC,BLDG_TYPE,LAND_SF,GROSS_AREA,BLDG_VALUE,LAND_VALUE,TOTAL_VALUE,GROSS_TAX,YR_BUILT,YR_REMODEL,ROOF_STRUCTURE,BED_RMS,FULL_BTH,KITCHENS,TT_RMS,RES_FLOOR,OVERALL_COND,HEAT_TYPE,AC_TYPE,PROP_VIEW,NUM_PARKING,FIREPLACES,ZIP_CODE))
```

## 3. Data Cleaning

* We'll clean the data using `IQR()` and replace the out of bound values by 'NA' and then impute 'NA' values by **Median** so that we can preserve the original sample size.

```{r}
clean_numeric_columns <- function(x) {
  # Remove outliers based on IQR
  q1 <- quantile(x, 0.25, na.rm = TRUE)
  q3 <- quantile(x, 0.75, na.rm = TRUE)
  iqr_value <- IQR(x, na.rm = TRUE)
  
  lower_bound <- q1 - 1.5 * iqr_value
  upper_bound <- q3 + 1.5 * iqr_value
  
  x_outliers_removed <- ifelse(x < lower_bound | x > upper_bound, NA, x)
  
  # Impute missing values with the median
  x_cleaned <- ifelse(is.na(x_outliers_removed), median(x, na.rm = TRUE), x_outliers_removed)
  
  return(x_cleaned)
}

# Clean all numeric columns in the dataset
property_subset_clean <- property_subset %>%
  mutate(across(where(is.numeric), clean_numeric_columns))

# Remove rows with any NA values after cleaning
property_subset_clean <- na.omit(property_subset_clean)

property_subset_clean <- property_subset_clean %>%
  mutate(CITY = ifelse(CITY == "" | is.na(CITY), "Unknown", CITY))

property_subset_clean <- property_subset_clean %>%
  mutate(OVERALL_COND = ifelse(OVERALL_COND == "" | is.na(OVERALL_COND), "Unknown", OVERALL_COND))

property_subset_clean <- property_subset_clean %>%
  mutate(PROP_VIEW = ifelse(PROP_VIEW == "" | is.na(PROP_VIEW), "Unknown", PROP_VIEW))

property_subset_clean <- property_subset_clean %>%
  mutate(AC_TYPE = ifelse(AC_TYPE == "" | is.na(AC_TYPE), "Unknown", AC_TYPE))
```

## 4. Tabulations
### a) Group statistics by City

* Grouping dataset by **cities** and generating statistics like **Mean**,  **sd**, **min**, **max**, and **count**

```{r}
city_stats <- property_subset_clean %>%
  group_by(CITY) %>%
  summarise(across(where(is.numeric), list(
    mean_value = ~mean(., na.rm = TRUE),
    sd_value = ~sd(., na.rm = TRUE),
    min_value = ~min(., na.rm = TRUE),
    max_value = ~max(., na.rm = TRUE),
    count = ~sum(!is.na(.))
  ), .names = "{col}_{fn}"))

kable(city_stats, format = "html",
      caption = "Grouping Statistics by CITY",
      align ="cccc",
     )%>%
  kable_styling(full_width = FALSE)

```

### b) Group statistics by overall condition of the parcel 

* Grouping dataset by **overall condition** and generating statistics like **Mean**,  **sd**, **min**, **max**, and **count**
```{r}
overall_cond_stats <- property_subset_clean %>%
  group_by(OVERALL_COND) %>%
  summarise(across(where(is.numeric), list(
    mean_value = ~mean(., na.rm = TRUE),
    sd_value = ~sd(., na.rm = TRUE),
    min_value = ~min(., na.rm = TRUE),
    max_value = ~max(., na.rm = TRUE),
    count = ~sum(!is.na(.))
  ), .names = "{col}_{fn}"))

kable(overall_cond_stats, format = "html",
      caption = "Grouping Statistics by OVERALL CONDITION",
      align ="cccc",
     )%>%
  kable_styling(full_width = FALSE)

```

### c) Group statistics by Property view

* Grouping dataset by **Property view** and generating statistics like **Mean**,  **sd**, **min**, **max**, and **count**
```{r}
prop_view_stats <- property_subset_clean %>%
  group_by(PROP_VIEW) %>%
  summarise(across(where(is.numeric), list(
    mean_value = ~mean(., na.rm = TRUE),
    sd_value = ~sd(., na.rm = TRUE),
    min_value = ~min(., na.rm = TRUE),
    max_value = ~max(., na.rm = TRUE),
    count = ~sum(!is.na(.))
  ), .names = "{col}_{fn}"))

kable(prop_view_stats, format = "html",
      caption = "Grouping Statistics by Property View",
      align ="cccc",
     )%>%
  kable_styling(full_width = FALSE)
```

### d) Statistics for grouping by AC Type 

* Grouping dataset by **Ac Type** and generating statistics like **Mean**,  **sd**, **min**, **max**, and **count**
```{r}
ac_type_stats <- property_subset_clean %>%
  group_by(AC_TYPE) %>%
  summarise(across(where(is.numeric), list(
    mean_value = ~mean(., na.rm = TRUE),
    sd_value = ~sd(., na.rm = TRUE),
    min_value = ~min(., na.rm = TRUE),
    max_value = ~max(., na.rm = TRUE),
    count = ~sum(!is.na(.))
  ), .names = "{col}_{fn}"))

kable(ac_type_stats, format = "html",
      caption = "Grouping Statistics by AC type",
      align ="cccc",
     )%>%
  kable_styling(full_width = FALSE)
```

### e) Statistics for grouping by LU 

* Grouping dataset by **LU** and generating statistics like **Mean**,  **sd**, **min**, **max**, and **count**
```{r}
LU_stats <- property_subset_clean %>%
  group_by(LU) %>%
  summarise(across(where(is.numeric), list(
    mean_value = ~mean(., na.rm = TRUE),
    sd_value = ~sd(., na.rm = TRUE),
    min_value = ~min(., na.rm = TRUE),
    max_value = ~max(., na.rm = TRUE),
    count = ~sum(!is.na(.))
  ), .names = "{col}_{fn}"))

kable(LU_stats, format = "html",
      caption = "Grouping Statistics by LU",
      align ="cccc",
     )%>%
  kable_styling(full_width = FALSE)
```

### f) Statistics for overall cleaned dataframe

* Generating statistics like **Mean**,  **sd**, **min**, **max**, and **count** for clean dataframe.
```{r}
numeric_columns <- c("LAND_SF", "GROSS_AREA", "BLDG_VALUE", "LAND_VALUE", "TOTAL_VALUE", 
                     "GROSS_TAX", "YR_BUILT", "YR_REMODEL", "BED_RMS", "FULL_BTH",
                     "KITCHENS", "TT_RMS", "RES_FLOOR", "NUM_PARKING", "FIREPLACES")

# Calculate summary statistics
summary_df <- map_dfr(numeric_columns, ~ {
  col_name <- as.symbol(.x)
  property_subset_clean %>%
    summarise(
      variable = .x,
      mean_value = mean(!!col_name, na.rm = TRUE),
      min_value = min(!!col_name, na.rm = TRUE),
      max_value = max(!!col_name, na.rm = TRUE),
      count = sum(!is.na(!!col_name)),
      sd_value = sd(!!col_name, na.rm = TRUE)
    )
})

summary_df
```

## 5. Sampling

### a) Creating 5 random samples with replacement using `sample_n()` and calculating descriptive statistics for each sample
```{r}
# Number of samples
num_samples <- 5

# Create a list to store sample summary data frames
sample_summary_list <- vector("list", length = num_samples)

# Create samples and calculate summary statistics
set.seed(123)  # for reproducibility
for (i in seq_len(num_samples)) {
  sample_data <- property_subset_clean %>%
  sample_n(size = nrow(property_subset_clean), replace = TRUE)
  
  sample_summary <- map_dfr(numeric_columns, ~ {
    col_name <- as.symbol(.x)
    sample_data %>%
      summarise(
        variable = .x,
        mean_value = format(mean(!!col_name, na.rm = TRUE),scientific=FALSE),
        min_value = min(!!col_name, na.rm = TRUE),
        max_value = max(!!col_name, na.rm = TRUE),
        count = sum(!is.na(!!col_name)),
        sd_value = format(sd(!!col_name, na.rm = TRUE),scientific =FALSE)
      )
  })
  
  sample_summary_list[[i]] <- sample_summary
}

# Combine the results into a single data frame
combined_summary_df <- bind_rows(sample_summary_list, .id = "Sample")

```


### b) Comparing the **Mean** values of **Gross Tax, Land Value and Land SF** for different samples

```{r}
# Columns of interest
columns_of_interest <- c("LAND_VALUE", "LAND_SF", "GROSS_TAX")

# Extract rows related to the columns of interest
mean_comparison_df <- combined_summary_df %>%
  filter(variable %in% columns_of_interest)

# Select and arrange relevant columns
mean_comparison_df <- mean_comparison_df %>%
  select("Sample", "variable", "mean_value") %>%
  spread(variable, mean_value)

# Print the mean comparison data frame
print(mean_comparison_df)
```
* Sample 1 has GROSS_TAX of **$6203.31**, LAND_SF of **2697.804**, and LAND_VALUE of **$95,624.71**.
* Sample 2 has GROSS_TAX of **$6186.348**, LAND_SF of **2688.85**, and LAND_VALUE of **$95,212.91**.
* Sample 3 has GROSS_TAX of **$6171.558**, LAND_SF of **2684.028**, and LAND_VALUE of **$94,983.86**.
* Sample 4 has GROSS_TAX of **$6189.36**, LAND_SF of **2693.415**, and LAND_VALUE of **$95,139.83**.
* Sample 5 has GROSS_TAX of **$6159.283**, LAND_SF of **2686.33**, and LAND_VALUE of **$94,431**.

### c) Comparing the 5 samples with **orginal** data for **Gross tax, Land sf and Land value**

```{r}
columns_of_interest <- c("LAND_VALUE", "LAND_SF", "GROSS_TAX")

# Extract rows related to the columns of interest from the original dataset summary
original_summary_df <- summary_df %>%
  filter(variable %in% columns_of_interest) %>%
  select(variable, 
         mean_value_original = mean_value,
         sd_value_original = sd_value)  # Include standard deviation

# Extract rows related to the columns of interest from the sampled data summary
sampled_summary_df <- combined_summary_df %>%
  filter(variable %in% columns_of_interest) %>%
  select("Sample", variable, 
         mean_value_sampled = mean_value,
         sd_value_sampled = sd_value)  # Include standard deviation

# Merge the two summary data frames based on the variable column
comparison_df <- merge(original_summary_df, sampled_summary_df, by = "variable")

# Print the comparison data frame
print(comparison_df)

```

### 1.GROSS TAX -
* The **mean** GROSS_TAX in the original dataset is approximately **$6187.05** with a **standard deviation** of **$4033.92**.
* Across five samples, the **mean** GROSS_TAX values range from **$6159.28 to $6203.31**, with corresponding **standard deviations** ranging from **$4016.11 to $4041.39**.

### 2.LAND_SF - 
* The **mean** LAND_SF in the original dataset is approximately **2689.26** with a **standard deviation** of **2162.20**.
* Across five samples, the **mean** LAND_SF values range from **2684.03 to 2697.80**, with corresponding **standard deviations** ranging from **2159.37 to 2165.25**.

### 3.LAND_VALUE - 
* The **mean** LAND_VALUE in the original dataset is approximately **$95,048.67** with a **standard deviation** of **$129,445.63**.
* Across five samples, the **mean** LAND_VALUE values range from **$94,431 to $95,624.71**, with corresponding **standard deviations** ranging from **$128,915.70 to $129,719.30**.

## 6. Visualisations

### a) Jitter Chart for Land value vs Property view

### Q. When is a jitter chart used?
* Jitter charts are useful when you have categorical or discrete data, and there is overlap among data points. They help avoid overplotting and provide a clearer view of the data distribution.

```{r}
jitter_prop_view_landvalue <- property_subset_clean %>%
  filter(PROP_VIEW != "Unknown") %>%
  select(LAND_VALUE, PROP_VIEW)

ggplot(jitter_prop_view_landvalue, aes(x = PROP_VIEW, y = LAND_VALUE)) +
  geom_jitter(width = 0.2, height = 0, color = "blue", size = 0.25) +
  labs(title = "Jitter Plot: LAND_VALUE vs. PROPERTY VIEW",
       x = "PROPERTY VIEW", y = "LAND_VALUE")+
    scale_y_continuous(labels = scales::comma) 
# Format y-axis labels with commas

```
* We plotted jitter chart using `ggplot()` and observed that our dataset consists of high values of **Average** rated properties followed by **Fair** and **Good**


### b) Scatter Plots

* We've used `par()` to display two scatter plots side-by-side and `plot()` to create the graphs
```{r}

# Set up the layout for multiple plots
par(mfrow = c(1, 2))  # 1x2 grid for multiple plots


# Change axis labels to non-scientific notation
options(scipen = 999)

# Scatter Plot: LAND_VALUE vs. LAND_SF
plot(property_dataset$LAND_SF, property_dataset$LAND_VALUE,
     main = "LAND_VALUE vs. LAND_SF",
     xlab = "LAND_SF", ylab = "LAND_VALUE", col = "blue",
     pch = 1,      # Change point type
     cex = 0.5)     # Change point size


# Scatter Plot: GROSS_TAX vs. GROSS_AREA
plot(property_dataset$GROSS_AREA, property_dataset$GROSS_TAX,
     main = "GROSS_TAX vs. GROSS_AREA",
     xlab = "GROSS_AREA", ylab = "GROSS_TAX", col = "blue",
     pch= 1,
     cex = 0.5
     )

options(scipen = 0)

# Reset the layout to default
par(mfrow = c(1, 1))

```
* The scatter plot on the left shows relation between Land_value and Land sf. We can notice outliers which are distant points away from main distribution.
* The scatter plot on the right shows relation between Gross tax and Gross area and a slight positive correaltion can be observed between the two variables. We can observe outliers which are distant from the main distribution.

### c) Boxplots

### Q. How can you use boxplots to detect outliers?

* In a boxplot, the box represents the interquartile range (IQR), which is the range between the first quartile (Q1) and   the third quartile (Q3).
* The median is represented by a line inside the box.
* "Whiskers" extend from the box to the minimum and maximum values within a certain range.
* Points beyond the whiskers are considered potential outliers.
* `abline()` is used commonly to detect or hihglight outliers by drawing a line at 1.5 times the IQR from the quartiles.



#### i) BOXPLOT for building value

```{r}

boxplot(property_subset_clean$BLDG_VALUE,
        main = "Boxplot: BLDG_VALUE",
        col = "lightblue",
        border = "darkblue",
        ylab = "BLDG_VALUE")

# Calculate the threshold for outliers (you can adjust the multiplier as needed)
threshold <- 1.5  # Adjust this value based on your requirements

# Add a horizontal line to the boxplot at the threshold with a red dashed line
abline(h = boxplot.stats(property_subset_clean$BLDG_VALUE)$out, col = "red", lty = 2, lwd = 0.5)

```

* The boxplot helps visually summarize and compare the central tendency, spread, and potential outliers in **Building values**. The red line indicates outliers.



#### ii) BOXPLOT for LAND VALUE by LU

* This plot allows us to visualize the distribution of land values within different categories of the LU variable

```{r}
options(scipen = 999)


# Adjust the margins to create more space for y-axis labels
par(mar = c(4, 5, 3, 2) + 0.5)  # c(bottom, left, top, right)

# Create the boxplot
boxplot(LAND_VALUE ~ LU, data = property_subset_clean,horizontal = TRUE,
        main = "Boxplot: LAND_VALUE by LU",
        col = "blue", ylab = "LU", las = 1)

lower_threshold <- 1.5  # Adjust this value based on your lower bound requirements
upper_threshold <- 10  # Adjust this value based on your upper bound requirements

abline(v = lower_threshold, col = "red", lty = 2)
abline(v = upper_threshold, col = "green", lty = 2)


# Reset the margins to default after creating the plot
par(mar = c(5, 4, 4, 2) + 0.1)  # Reset to default values

options(scipen = 0)

```
* The boxplot above shows the distribution of Land values within LU's. The values beyond the whiskers are **potential outliers**. 
* Majority of the LU's have right skew since their right whiskers are longer.

#### iii) BOXPLOT for LAND VALUE by AC Type

* This plot allows us to visualize the distribution of land values within different categories of the AC_TYPE variable
* Plotted graph using `boxplot()` and also used `abline()` for outlier detection.
```{r}
options(scipen = 999)

boxplot(LAND_VALUE ~ AC_TYPE, data = property_subset_clean,
        main = "Boxplot: LAND_VALUE by AC_TYPE",
        col = "lightgreen",
        border = "darkgreen",
        xlab = "AC_TYPE",
        ylab = "LAND_VALUE")

# Calculate the threshold for outliers (you can adjust the multiplier as needed)
threshold <- 1.5  # Adjust this value based on your requirements

# Add a horizontal line to the boxplot at the threshold
abline(h = boxplot.stats(property_subset_clean$LAND_VALUE)$out, col = "red", lty = 2, lwd = 0.5)

options(scipen = 0)

```
* The boxplot helps visually summarize and compare the central tendency, spread, and potential outliers in land values for each "AC_TYPE" category.
* The Land value within **Central AC** is more skewed than **Ductless AC**.


#### iv) BOXPLOT for Gross Area by Heat Type

* This plot allows us to visualize the distribution of gross area within different categories of the HEAT_TYPE variable


```{r}

# Adjust the margins to create more space for y-axis labels
par(mar = c(4, 8, 3, 2) + 0.5)  # c(bottom, left, top, right)

boxplot(GROSS_AREA ~ HEAT_TYPE, data = property_subset_clean,horizontal = TRUE,
        main = "Boxplot: GROSS_AREA by HEAT_TYPE",
        col = "lightcoral",
        border = "darkred",
        xlab = "GROSS_AREA",
        ylab = "",
        names.arg = c("Electric", "Forced", "Label3", "Label4"),  # Replace with your actual labels
        las = 2)  # Rotate x-axis labels at a 45-degree angle

# Reset the margins to default after creating the plot
par(mar = c(5, 4, 4, 2) + 0.1)  # Reset to default values
```
* The boxplot helps visually summarize and compare the central tendency, spread, and potential outliers in **Gross area** for each **Heat_Type** category.













